Comprehensive optical and data management infrastructure for high-throughput light-sheet microscopy of whole mouse brains

### 2015-03-25 Neurophotonics/latex-style-files-spie-manuscript/manuscript1.1_Ludo.tex
commit

The stack component is used to acquire single image stacks along the $z$ direction, and to this aim send specific messages to coordinate the Camera and the Stages components. Stacks are collected by moving continuously the sample along $z$ while keeping the camera acquiring in free run. The translation speed is such that in the time needed to collect a single frame the specimen has moved by a $z$ step specified in the GUI. The stack component also specify the name of the file where data are saved.

@@

The volume of a mouse brain have been estimated from MRI measurements \cite{Kovacevic2005} as $\simeq 0.5$ $\text{cm}^3$; the same Authors find out that a parallelepiped encompassing the whole brain would have a volume of $\simeq 0.9$ $\text{cm}^3$. Since the clearing procedure with CLARITY produces a linear expansion of about $1.2\times$ \cite{Costantini}, the volume of a clarified mouse brain can be estimated as $\simeq 0.9$ $\text{cm}^3$, while that of a parallelepiped encompassing the sample is expanded to $\simeq 1.6$ $\text{cm}^3$. Given the pixel size of the camera and the magnification of the system, and assuming a $z$ step of 2 $\mu$m, the voxel size results $0.26\times0.26\times2$ $\mu\text{m}^3$. This means that a cleared mouse brain will be represented with about 6 TeraVoxels, while a parallelepiped encompassing the brain will result in 10.8 TeraVoxels. Since the camera produces 16 bit images with no compression, each raw voxel occupies 2 bytes of disk space, resulting in 12 TB for the whole brain "alone" and 21.6 TB for the parallelepiped (see Table XXXX).

@@

The camera used in our light sheet microscope (Hamamatsu Orca Flash 4.0 v2) flushes the datastream directly to a dedicated workstation with 12 solid-state drives (SSD) operating in RAID 0 configuration, resulting in a virtual drive of 10.9 TB. Although this is already a large and expensive system for the current SSD technology, it is way smaller than the size of the raw datasets produced in each tomography. We thus connected the workstation to a larger network area storage (NAS) via a 10 Gbit/s network copper cable. The NAS operates in a safer data modality (RAID 5) with 32 TB of useful disk space. The NAS is then connected by a dedicated 10 Gbit/s link provided by the Consortium GARR (the Public Institution managing research networks in Italy) to CINECA (the main italian supercomputing center).

The camera used in our light sheet microscope has a dedicated work station with a CameraLink card and 10.9TB of SSD space in RAID 0 configuration. This finite storage capacity is further expanded by a network-attached storage (NAS) of 32.4TB (Figure\ref{fig:DataFlow}). The SSDs and the NAS are connected via a 10Gb bandwidth connection using a 10Gb switch and a 10Gb optical fibre connects our lab to the supercomputing centre CINECA in Bologna. Mention PICO system/storage analysis capacity there? Mention graphical processing unit and cores/processors/RAM of work station?

With the size of a single image of 8MB and a stack made up of say 3900 frames a single stack amounts to 30.4GB of disc space (see Table\ref{tab:dataproduction} for an overview of the values). Based on a frame rate of 44Hz each stack takes approximately 1.5 minutes to acquire and the SSD RAID fills up with a data writing rate of 1.2TB/h. If no further measures are taken the SSD RAID of 10.9TB would be filled up in just 9h. A custom compression script written in Python serves the triple purpose of translating the images from the proprietary Hamamatsu file format to 3D tif, to transfer the data from the SSD RAID to a network attached storage (NAS) and to reduce the raw data size by first, digitally downsampling the image resolution and second, applying a lossless compression to the tif images. This leads to a overall reduction of data size by a factor of ten, each stack now taking up only 3GB of disc space, albeit at a much slower conversion rate (-0.7TB/h) than the rate at which data is written to the SSD thus representing the most serious bottleneck to our data production pipeline. The SSD RAID fills up at a rate of 0.5TB/h which means the total acquisition time is limited to less than 24h and a total tomo size of just over 25TB. After this time acquisition has to be halted until the slower conversion process has caught up and freed the SSD RAID before data acquisition can start again. A whole mouse brain tomography would take 3 days of uninterrupted measuring yet with the necessary breaks for data transfer this time is easily doubled to tripled.

>>>>

The stack component is used to acquire single image stacks along the $z$ direction, and to this aim send specific messages to coordinate the Camera and the Stages components. Stacks are collected by moving continuously the sample along $z$ while keeping the camera acquiring in free run. The translation speed is such that in the time needed to collect a single frame the specimen has moved by a $z$ step specified in the GUI. The Camera component uses low-level libraries from Hamamatsu to save the data, producing one single \emph{.cxd} file (which is a proprietary tiff-like format from Hamamatsu) for each image stack. The name of this file and the path to it is specified by the Stack component.

@@

The volume of a mouse brain have been estimated from MRI measurements \cite{Kovacevic2005} as $\simeq 0.5$ $\text{cm}^3$; the same Authors find out that a parallelepiped encompassing the whole brain would have a volume of $\simeq 0.9$ $\text{cm}^3$. Since the clearing procedure with CLARITY produces a linear expansion of about $1.2\times$ \cite{Costantini}, the volume of a clarified mouse brain can be estimated as $\simeq 0.9$ $\text{cm}^3$, while that of a parallelepiped encompassing the sample is expanded to $\simeq 1.6$ $\text{cm}^3$. Given the pixel size of the camera and the magnification of the system, and assuming a $z$ step of 2 $\mu$m, the voxel size results $0.234\times0.234\times2$ $\mu\text{m}^3$. This means that a cleared mouse brain will be represented with about 7.4 TeraVoxels, while a parallelepiped encompassing the brain will result in 13.3 TeraVoxels. Since the camera produces 16 bit images with no compression, each raw voxel occupies 2 bytes of disk space, resulting in 14.8 TB for the whole brain "alone" and 26.6 TB for the parallelepiped (see Table XXXX). However, image stacks are acquired with a 10\% of linear superposition, which is used later for image stitching. This redundancy expands the size of raw data to 18.3 TB and 32.8 TB, respectively.

@@

The camera used in our light sheet microscope (Hamamatsu Orca Flash 4.0 v2) flushes the datastream directly to a dedicated workstation with 12 solid-state drives (SSD) operating in RAID 0 configuration, resulting in a virtual drive of 10.9 TB. Although this is already a large and expensive system for the current SSD technology, it is way smaller than the size of the raw datasets produced in each tomography. We thus connected the workstation to a larger network area storage (NAS) via a 10 Gbit/s network copper cable. The NAS operates in a safer data modality (RAID 5) with 32.4 TB of useful disk space. The NAS is then connected by a dedicated 10 Gbit/s link provided by the Consortium GARR (the Public Institution managing research networks in Italy) to CINECA (the main italian supercomputing center). The data flow scheme is summarized in Figure \ref{fig:DataFlow}.

While transferring the data from the SSD to the NAS, we also perform a file conversion from the proprietary Hamamatsu format to a more manageable one. We chose the TIFF format among the others because it permits to store 3D image with lossless compression in a single file. Furthermore, it is a flexible and platform-independent format supported by numerous image processing applications. We thus developed a conversion module that leverages the functionalities offered by the interface of the DCIMGAPI library, supplied by Hamamatsu, and the freely available LibTIFF library. The module reads a single .cxd 3D image and writes it in a variable number of multi-page TIFF files, depending on the user-defined limit of pages per file. Since the output of the Hamamatsu camera may exceed tens of Gbytes, the image in proprietary format is read slice by slice to minimize memory requirements. The module has also an option to downsample the image on save the input image at full or reduced resolution, in order to adapt to storage availability. Usually we downsample the image in the XY plane by a factor 2 and leave the Z sampling as usual. This kind of downsampling, in combination with the lossless compression offered by the TIFF format, reduces data size by about one order of magnitude. Thus the parallelepid volume cointaining the whole mouse brain can be represented using about 2-3 TB of disk space.

Although file conversion drastically reduces data size, in the current implementation it is quite slower than data production. The Camera component flushes data on the SSD with a rate $R_{data}$ dependent on frame rate $R_{frame}$ and on the image size:
\begin{equation}
R_{data} [\text{Bytes}] = 2\times R_{frame} \times \#_{pixels}\,,
\end{equation}
where the factor takes into account the 16 bit grey depth of images. The volumetric imaging rate is given by $R_{volume} = 0.5 R_{data} \times \text{Voxel size}$.

With a standard image size of $2048\times 2048$ pixels, and $R_{frame}=44$ Hz, data are produced at a rate $R_{data}=352$ MB/s $\approx 1.2$ TB/h, which results in a volumetric rate $R_{volume} \approx 2\times 10^7$ $\mu\text{m}^3/s$ $\approx 0.07$ $\text{cm}^3$/h. The conversion module can empty the SSD wih an average rate of about 0.7 TB/h, meaning that the SSD is actually filled up during imaging at a rate of about 0.5 TB/h. Imaging sessions should thus be shorter than 21.8 hours: this limits the maximum amount of raw data being acquired continuously to approximately 25.4 TB. Taking into accont the 10\% linear superposition between adjacent stacks, the maximum size of a sample that can be imaged in a single session is about 1.2 $\text{cm}^3$, which is larger than a cleared mouse brain, but smaller than the estimate size of a parallelepiped volume encompassing the entire volume. Therefore imaging should be limited to the portion of the volume where the brain is with rational selection of stacks. Larger specimens, as rat brain or portion of human brain, must be imaged in consecutive session until faster strategies for image conversion are devised.

### 2015-03-25 Neurophotonics/latex-style-files-spie-manuscript/manuscript_snellendo_Ludo.tex
commit

%\bibitem{Frasconi2014} P. Frasconi, L. Silvestri, P. Soda, R. L. Cortini, F. S. Pavone and G. Iannello, ``" \emph{Bioinformatics} \textbf{30}(), i587-i593 (2014).

@@

\subsubsection{Sample preparation for whole brain imaging}

>>>>

\bibitem{Frasconi2014} P. Frasconi, L. Silvestri, P. Soda, R. L. Cortini, F. S. Pavone and G. Iannello, "Large-scale automated identification of mouse brain cells in confocal light sheet microscopy images" \emph{Bioinformatics} \textbf{30}(17), i587-i593 (2014).

@@

\subsubsection{Sample preparation for whole brain imaging}
Mouse brains were processed using the CLARITY protocol \cite{Chung2013} and cleared with a new refractive-index-matching medium \cite{Costantini}. Adult mice (post-natal day 56) were anaesthetized with isoflurane and transcardially perfused with 20 ml ice-cold phosphate-buffered saline (PBS) solution (pH 7.6) followed by 20 ml of a mixture of 4\% (wt/vol) paraformaldehyde (PFA), 4\% (wt/vol) acrylamide, 0.05\% (wt/vol) bis-acrylamide, 0.25\% (wt/vol) VA044 in PBS. We then extracted the brains from the skull and put them in the same PFA/acrylamide/bis-acrylamide/VA044/PBS solution at 4°C for 3 days. Afterwards, we degassed the samples and increased their temperature to 37 °C to initiate polymerization. The embedded samples were then extracted from the gel and washed with clearing solution solution (sodium borate buffer (200 mM, pH 8.5) containing 4\% (wt/vol) sodium dodecyl sulfate) at 37 °C through gentle shaking. To perform electroforetic tissue clearing (ETC), hydrogel-embedded brains were placed in a custom-built organ-electrophoresis chamber. The same clearing solution was circulated through the chamber and a voltage of 20V was applied across the ETC chamber at 37 °C for several days. After clearing, brains were incubated in PBST0.1 (PBS and 0.1\% Triton X-100, pH 7.6) at 37 °C for 2 days to remove the clearing solution. As a final step, murine brain were incubated in 50 ml of 30\% and 63\% (vol/vol) 2,2'-thiodiethanol in 0.01M PBS (TDE/PBS), each for 1 day respectively at 37°C while gently shaking. The final 63\% TDE/PBS solution has a refractive index $n\simeq 1.45$, allowing for refractive index matching and thus tissue transparency.
